Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-07-04 main_revision.tex
Updates from Overleaf

\usepackage{}\documentclass[draft,linenumbers]{agujournal}

@@

\emph{froot\_leaf} had contrasting impacts at different sites.  At MET,  where low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions. At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.

@@

\subsection{Low leaf area investigations}
In general, we found that leaf area indices predicted by the CLM5 default parameter set were often low  compared to published estimates for these sites. (for Caxiuan\~a, 3.0 (modeled) vs. 5.3 (observed) \cite{fisher2007}, for Niwot Ridge, 1.7 vs. 4.2 \cite{bowling2009}, for Harvard Forest 2.6 vs. 3.5 \cite{williams1996} and for Metolius, 0.5 vs. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran an entire separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations. For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs. 1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs. 1.1, 1.26, 1.21 for the control). Similar results were found at all other sites.  Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology.

\section{Discussion}
The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.

At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (CLM4 and CLM4.5, Wieder et al. 2018). Given the dominant impact of N fixation on the model fertilization responses, it seems likely that the capacity for the model to fix addition N in response to CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.  Intuitively, it seems likely that an ecosystem that could more rapidly shift a large fraction of resources to N uptake would

The CLM5 model has a number of new prognostic elements, which allow predictions in elements of the C and N cycles which were assumed to be constant in previous versions of the model. Exploration of the parameter space that controls these new elements allows a clear exploration of the sensitivities of the output space of the model to critical assumptions.

The model output space is notable complex, and in many cases the sign of the impact of the different parameters changes between sites, and trade-offs in the influence of carbon and nitrogen economic parameters interact with both the initial state and environmental forcing. That the parameter space of CLM5 is complex is not surprising, given its relatively high process fidelity, but a) this type of sensitivity analysis,  illustrating how different parameterizations respond to different growth conditions and environmental forcing, is relatively rare for models of this kind,  b) the large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and c) the results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites. Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.

Further, our results indicate that, in line with our initial hypothesis 2, model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization, since for each site there was substantial lack of overlap between the parameters controlling the default state and the fertilization responses.  We did observe, however, many instances where simulations with high leaf area index responded less strongly to fertilization, and thus the numerous parameters impacting this feature of the initial model state (dominated by tissue allocation ratios and \emph{slatop}) would also have an indirect impact on fertilization.  Our results suggest a more general point, that high LAI systems should respond less to fertilization. We did not find a clear statistical link between initial LAI and model output across the whole parameter space, but interpretation of global patterns in terrestrial cardon storage should likely take this mechanism into account.

@@

\subsection{The status of the CLM5 representation of plant carbon and nitrogen representation}
Our efforts to improve the fidelity of the nitrogen cycle in the CLM5 were primarily motivated by the need to move the model behaviour closer to several key dynamic features of the system. Notably, the dynamics of $N_{leaf}$ with respect to environmental scarcity (\cite{lovelock1998}, \cite{brzostek2014}), the dynamics of photosynthetic capacity with respect to environmental conditions (\cite{xu2012}, \cite{ali2016},\cite{rogers2017}, \cite{bloomfield2018}), the changes in N fixation rate with vegetation state and the scarcity of N (\cite{vitousek2002}), and to include a means of deploying carbon as a means to acquire additional nitrogen (\cite{terrer2018}). The latter acts as a proxy for the numerous ways in which plant symbiont systems might adapt to low nutrient conditions, including increases in fine root exudation and allocation, fungal biomass, tissue metabolic rates and turnover/nutrient foraging.  In doing so, we integrate a suite of physiological processes that lack complete mechanistic and quantitative understanding.  By investigating the critical roles of these processes in the model dynamics, we hope to highlight a research need, and to stimulate research on and improvements in the representation of these processes.

In analyzing the results of this study, for example, we found numerous instances where the model reduced carbon use efficiency (CUE, defined here as net primary productivity, NPP, divided by gross primary productivity, GPP, where the respiratory carbon used for N uptake is not counted within NPP and thus lowers CUE) by changing tissue C:N ratios in response to forcing. In principle, there exists an optimal leaf C:N ratio that maximizes leaf C export under a given set of conditions. Despite its optimal allocation of leaf N \emph{within} the leaf (to photosynthetic enzymes) via the LUNA model, CLM5 does not include any scheme that ensures that overall tissue CN ratios conform to any physiological optima. Instead, a heuristic model adjusts tissue CN ratios round the target ratio (an input parameter).  Numerous models of this type have been proposed in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{anten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}). We chose not to include this feature in CLM5 on account of limited evidence for its statistical strength in analyses of the global leaf trait datasets used in  LUNA, but future nitrogen cycling model development efforts could re-visit this topic.

Another area requiring focus is the dominance of the 'fraction of fixers' over fertilization responses. \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation. Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}). The 0-1 range of \emph{frac\_fixers} used here is ecologically unlikely, given the costs borne by plants that can fix N give them a competitive disadvantage in systems no longer limited by N. However, it suggest that a high CO$_{2}$ world would likely provide much greater benefits to N fixation, and that those benefits would be ameliorated by ongoing increases in N deposition.

\subsection{ Methodological Limitations}
\subsubsection{Limitations of step-change experiments}
Here we investigate the parameter space of the model using step-change experiments. These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions. Further, we expect that the carbon and nitrogen cycles may take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.

To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments. At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, (result not shown) to those at 15 years.  However, at 5 years into the fertilization in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent (a x~2.06 impact on NPP, compared to ~0.95 in the year 15 analysis at Niwot Ridge, indicating that the results described above had already been subject to progressive N limitation, and were closer to an equilibrated state.

@@

\subsubsection{Temperature and water responses}
In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model to transient climate forcing has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.

>>>>

\documentclass[draft,linenumbers]{agujournal}

@@

\emph{froot\_leaf} had contrasting impacts at different sites.  At MET,  low values increased the CO$_{2}$ fertilization, probably due to the increased drought stress under those conditions. At HVF, low \emph{froot\_leaf} decreased both CO$_{2}$ and N fertilization on account of higher initial LAI, mirroring the respnse to \emph{slatop}.

@@

\subsection{Low leaf area investigations}
In general, we found that LAI predicted by the default CLM5 were often low  compared to published estimates for these sites. (CAX; 3.0 (modeled) vs. 5.3 (observed) \cite{fisher2007},  NWR; 1.7 vs. 4.2 \cite{bowling2009}, HVF; 2.6 vs. 3.5 \cite{williams1996}  MET; 0.5 vs. 1.5 (\cite{spadavecchia2011}). This is a somewhat surprising result, given the generally good agreement of globally-driven model output with LAI reported by \cite{lawrence2018}, and may plausibly reflect differences between the use of reanalysis products used to drive the model and site-specific meteorological data.

To interrogate this further, we ran a separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher (and more consistent with observed) LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. For example, at HVF, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response of  (1.18, vs 1.43 for the control)  and a lower N response as well (1.20 vs 1.26 for the control). Similar results were found at all other sites.

\section{Discussion}
The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space. Interpretation of these simulations should be informed by solid comprehension of which model properties dominate the size of the feedback strength both within and between models, and we intend these illustrative simulations to provide initial guidance of this process.

At the time of writing, we do not yet know the carbon-climate feedback strength of the default version of CLM5/CESM2. In offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ and a weaker response to N fertilization than previous versions of the model (CLM4 and CLM4.5, Wieder et al. 2018). Further, the impact of both types of fertilization was found, in all versions of the model, to be highly variable with PFT (from 0\% to 110\% for N, -10\% to +55\% for CO$_{2}$). In this study we illustrate how variations in the pre-fertilization state, water and N limitation, and parametric variation combine to generate variation in the responses to forcing. While we set out primarily to to determine responses to model parameters, findings concerning the impact of model state on forcing sensitivity will have important implications for the interpretation of simulations with the default model.

Given the dominant impact of N fixation (\emph{fracfixers}) on the model fertilization responses, it seems likely that the new capacity, introduced in CLM5, for plants to use carbon to pay for the fixation of addition nitrogen under CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.

Aside from the straightforward direct impacts of the fixation rate on model predictions, however, the model output space was notably complex. For example,  in some cases the sign of the impact of individual parameters changes between sites, while other parameters can demonstrate non-linear and counter-intuitive impacts as the result of numerous interacting feedbacks.  To summarize our findings, we note a variety of key outcomes of our analysis on the factors controlling responses to CO$_{2}$ and N fertilization in the CLM5, including:

1. The initial model state has a dominant effect on  responses to CO$_{2}$ and N fertilization. The initial leaf area index, and whether the system is on a linear or saturated part of the LAI vs GPP relationship, has a large impact on any fertilization response. Parameter impacts in more productive systems are likely to generate smaller LAI/GPP feedbacks and lower parametric variation as a result. Many of the direct impacts of the parameters assessed here were masked by their large impacts on the initial model state.  The clearest examples of this were the \emph{slatop} and \emph{stem\_leaf} impacts, but this impact was illustrated for many model configurations.

2. The degree of water or nutrient limitation in the initial state is also a dominant factor over the model responses to CO$_{2}$ fertilization and N deposition. Once again, the direct impact of parameter perturbation is frequently masked by how changes affect the initial limiting conditions. Examples of this include how low \emph{froot\_leaf} increases water limitation at NWR, increasing responsiveness to CO$_{2}$ and how high \emph{leafcn} reduces water stress and thus decreases the CO$_{2}$ response at CAX.

3. Other parameters with a direct impact on fertilization responses included those related to the acquisition of nitrogen (\emph{N\_costs}) and the stoichiometric flexibility of the system (\emph{cn\_flex\_c}) illustrate the sensitivity of the model to relatively unconstrained processes within the nitrogen cycle.

4. Many of the parameters that are dominant over the initial state (\emph{slatop}, \emph{stem\_leaf} \emph{leafcn}) either have limited impacts on fertilization responses, or have impacts that are largely traceable to their impacts on the initial model state. In contrast, numerous parameters that more directly and uniformly impact fertilization responses, (\emph{frac\_fixers} \emph{N\_costs}) are less important for the initial model state. We thus suggest that model calibration efforts that only take into account parameters that affect the control model state, are unlikely to constrain the major axes of responses to CO$_{2}$ and N fertilization.

5. The results can in general be interpreted as being in line with the expected dynamics of the model, given sufficient output data and consideration of the dynamics of the individual sites. Given the complexity of the model, this helps generate confidence that the model outcomes are derived from the intended process mechanisms, and are not subject to significant deviations from the intent of the model structure (as opposed to .

6. The large degree of variability in the responses between sites and output variables will likely hinder attempts to generate land surface models with globally optimized parameters, and this analysis should help inform and guide the design of such efforts and  Sensitivity tests across broader geographical ranges would help to eludicate more clearly how changes in baseline water and N limitation impact parameter responses and patterns of responses to fertilization.

\subsection{Applicability of results to other land surface models}
Our finding that Nitrogen fixation plays a dominant role over responses to $CO_{2}$ fertilization concurs with earlier studies. \cite{wieder2015} investigated the role of two alternative linear models of N fixation within the CLM4.5, while \cite{meyerholt2015} investigate a range of alternative models of N fixation implemented within the O-CN model (\cite{zaehle2010}.

@@

\subsection{Model improvements highlighted by this analysis}
While there are many potential improvements that could be made to any land surface model, this particular assessment of the parameter space of the CLM5 in this study revealed several dynamics which highlight parts of the representation of biogeochemistry which might be improved in future versions.  Firstly, the response of NPP to increasing \emph{slatop} was often negative, despite increasing GPP, as the fixed-fraction allocation scheme in the model will generate leaves that respire more than they assimilate under very high productivity conditions. This highlights the need for either more realistic tissue allocation, or for a scheme which optimizes leaf allocation to prevent leaf layers in negative carbon balance (\cite{fisher2010}).  Secondly, in other cases, the heuristic C:N stoichiometry algorithm can increase tissue N content in response to environmental forcing in a manner which, again, can act to reduce net assimilation by modifying the balance between assimilation and respiration. In principle, there exists an optimal leaf C:N ratio that maximizes leaf C export under a given set of conditions.  Numerous models of this type have been proposed in the plant optimal functioning literature (\cite{vanwijk2003}, \cite{mcmurtrie2011}, \cite{anten2011} \cite{franklin2012}, \cite{mcmurtrie2013}, \cite{thomas2014}) and future nitrogen cycling model development efforts could re-visit this topic.

Another area requiring focus is the dominance of the 'fraction of fixers' over fertilization responses. \emph{frac\_fixers} is a proxy for the community level composition of nitrogen fixing plants. Technically, because plant density is not represented in the CLM5, this parameter represents the fraction of net assimilated carbon ($gpp$ - $r_{auto}$) that can be used for symbiotic N fixation. Ideally, this feature of the model would be prognostic, and the fraction of N fixing plants would respond to changes in environmental conditions and the relative competitive abilities of N fixing plants. In general, Nitrogen fixers are known to primarily exist in early-successional parts of ecosystems (\cite{vitousek1989}), and therefore representation of their ecological dynamics would likely include a representation of vegetation demographics and succession (\cite{fisher2018vegetation}, \cite{trugman2016climate}).

\subsection{ Methodological Limitations}
Here we investigate the parameter space of the model using step-change experiments. These numerical experiments have the advantage that they can be informed by field experiments (\cite{wieder2019}, however, such experiments necessarily involve unrealistically rapid changes in environmental boundary conditions, and cannot necessarily be considered as analogous to either real-world or transient future scenario conditions.

@@

In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model to transient climate forcing has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters that in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implicitly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations.
